Visual diagnostics for constrained optimisation with application to guided tours

### 2020-10-08 paper/paper.Rmd
include the plots for comparing diff seeds

p <- bind_rows(holes_1d_geo, holes_1d_better) %>% 
  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
                   index = tourr::holes(), raw_data = data) %>% 
  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
                   index = tourr::holes(), raw_data  = data)%>%
  explore_space_pca(col = method)  + 
  scale_color_botanical(palette = "cherry") + 
  theme(legend.position = "bottom")

@@

```{r}
set.seed(123)
seed <- sample.int(10000, 20)
# kol_1d_2var_better_sim <- list()
# kol_1d_2var_better_sim_polish <- list()
# for (i in 1: length(seed)){
#   set.seed(seed[i])
#   kol_1d_2var_better_sim[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_better, max.tries = 100),
#                  rescale = FALSE, verbose = TRUE)
# 
# 
#   last_basis <- get_best(kol_1d_2var_better_sim[[i]])$basis %>% .[[1]]
# 
#   set.seed(1842)
#   kol_1d_2var_better_sim_polish[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_polish, max.tries = 100),
#                  rescale = FALSE, verbose = TRUE,
#                  start = last_basis)
# }

@@

polish_list <- purrr::map2(kol_1d_2var_better_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
better_list <- purrr::map2(kol_1d_2var_better_sim,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
polish_tidy <- do.call(dplyr::bind_rows, polish_list)
better_tidy <- do.call(dplyr::bind_rows, better_list)
all_tidy <- bind_rows(better_tidy, polish_tidy)
```

```{r seed-table, eval = FALSE}
best <- get_best(polish_tidy, group = seed)

all_summary <- best$basis %>% 
  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
  as_tibble() %>% 
  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
  mutate(IndexVal = best$index_val, seed = as.factor(seed)) %>% 
  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
  group_by(seed, IndexVal) %>% 
  filter(abs(value) == max(abs(value))) %>% 
  arrange(IndexVal) %>% ungroup()

all_summary %>% knitr::kable(digits = 3, caption = "this is the caption", format='latex', booktabs=TRUE, escape=FALSE)

```{r eval = FALSE}
pca <- all_tidy %>% 
  mutate(seed = as.factor(seed)) %>% 
  left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
                                        index = tourr::norm_kol(),
                                        raw_data = boa6) %>% 
   bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
                                        index = tourr::norm_kol(),
                                        raw_data = boa6) %>%

@@

pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
  group_by(seed) %>% arrange(IndexVal) %>% 
  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
  geom_point(data = random, col = "grey") +
  geom_point(aes(col = VarFound), alpha = 0.5) + 
  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
  geom_point(data = pca$aug %>% get_best(group = seed), size = 7) +
  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
  #theme_void() + 
  theme(aspect.ratio = 1) + 
  scale_color_botanical(palette = "fern")  
```

```{r eval = FALSE}
set.seed(123)
seed <- sample.int(10000, 20)
# kol_1d_2var_better_sim <- list()
# kol_1d_2var_better_sim_polish <- list()
# for (i in 1: length(seed)){
#   set.seed(seed[i])
#   kol_1d_2var_better_sim[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_better, max.tries = 100),
#                  rescale = FALSE, verbose = TRUE)
# 
# 
#   last_basis <- get_best(kol_1d_2var_better_sim[[i]])$basis %>% .[[1]]
# 
#   set.seed(1842)
#   kol_1d_2var_better_sim_polish[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_polish, max.tries = 100),
#                  rescale = FALSE, verbose = TRUE,
#                  start = last_basis)
# }

@@

all_summary <- best$basis %>% 
  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
  as_tibble() %>% 
  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
  mutate(IndexVal = best$index_val, seed = as.factor(seed)) %>% 
  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
  group_by(seed, IndexVal) %>% 
  filter(abs(value) == max(abs(value))) %>% 
  arrange(IndexVal) %>% ungroup()

all_tidy %>% mutate(seed = as.factor(seed)) %>% 
  left_join(all_summary, by = "seed") %>% 
  explore_trace_interp(group = seed, col = VarFound) + 
  facet_wrap(vars(fct_reorder(seed, IndexVal)))  + 
  #geom_hline(yintercept = theo1, alpha = 0.5, linetype = 2) + 
  #geom_hline(yintercept = theo2, alpha = 0.5, linetype = 2) + 
  scale_color_botanical(palette = "banksia")  + 
  theme( aspect.ratio = 1)

@@

pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
  group_by(seed) %>% arrange(IndexVal) %>% 
  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
  geom_point(data = random, col = "grey") +
  geom_point(aes(col = VarFound), alpha = 0.1) + 
  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
  geom_point(data = pca$aug %>% get_best(group = seed), size = 7) +
  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
  #theme_void() + 
  theme(aspect.ratio = 1) + 
  scale_color_botanical(palette = "fern") 
```

@@

dt <- bind_rows(all_tidy %>% filter(seed == 1842) %>% mutate(algo = "better V2"),
          all_tidy %>% filter(seed == 9982) %>% mutate(algo = "better V7"),
          kol_1d_2var_better_random %>% mutate(algo = "better random"))

@@

p2 <- pca$aug %>% filter(method != "randomly_generated" & info != "theoretical") %>% 
  ggplot(aes(x = PC1, y = PC2, col = algo)) + 
  geom_point(data = random, col = "grey") +
  geom_point(aes(col = algo), alpha = 0.1) + 
  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
  geom_point(data = pca$aug %>% get_best(group = algo), size = 5) +
  geom_point(data = pca$aug %>% get_start(),  size = 3) +
  facet_wrap(vars(algo)) +
  theme_void() + 
  theme(legend.position = "none", aspect.ratio = 1) + 
  scale_color_botanical(palette = "banksia")

>>>>

p <- bind_rows(holes_1d_geo, holes_1d_better) %>% 
  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
                   index = tourr::holes(), raw_data = boa5) %>% 
  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
                   index = tourr::holes(), raw_data  = boa5)%>%
  explore_space_pca(col = method)  + 
  scale_color_botanical(palette = "cherry") + 
  theme(legend.position = "bottom")

@@

```{r fig.height = 8}
set.seed(123)
seed <- sample.int(10000, 20)
# kol_1d_2var_better_sim <- list()
# kol_1d_2var_better_sim_polish <- list()
# for (i in 1: length(seed)){
#   set.seed(seed[i])
#   kol_1d_2var_better_sim[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_better, max.tries = 100),
#                  rescale = FALSE, verbose = TRUE)
# 
# 
#   last_basis <- get_best(kol_1d_2var_better_sim[[i]])$basis %>% .[[1]]
# 
#   set.seed(1842)
#   kol_1d_2var_better_sim_polish[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_polish, max.tries = 100),
#                  rescale = FALSE, verbose = TRUE,
#                  start = last_basis)
# }

@@

polish_list <- purrr::map2(kol_1d_2var_better_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
better_list <- purrr::map2(kol_1d_2var_better_sim,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
polish_tidy <- do.call(dplyr::bind_rows, polish_list)
better_tidy <- do.call(dplyr::bind_rows, better_list)
all_tidy_better <- bind_rows(better_tidy, polish_tidy)

best <- get_best(all_tidy_better, group = seed) %>% arrange(index_val)

all_summary <- best$basis %>% 
  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
  as_tibble() %>% 
  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
  mutate(IndexVal = best$index_val, seed = as.factor(best$seed)) %>% 
  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
  group_by(seed, IndexVal) %>% 
  filter(abs(value) == max(abs(value))) %>% 
  arrange(IndexVal) %>% ungroup()

pca <- all_tidy_better%>% 
  mutate(seed = as.factor(seed)) %>% 
  left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
                                        index = tourr::norm_kol(),
                                        raw_data = boa6) %>% 
   bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
                                        index = tourr::norm_kol(),
                                        raw_data = boa6) %>%

@@

pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
  group_by(seed) %>% arrange(IndexVal) %>% 
  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
  geom_point(data = random, col = "grey", size = 0.3) +
  geom_point(aes(col = VarFound), alpha = 0.1) + 
  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
  geom_point(data = pca$aug %>% get_best(group = seed), size = 5) +
  geom_point(data = pca$aug %>% get_start() %>% filter(method == "search_better_random"), size = 3) +
  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
  theme_void() + 
  theme(aspect.ratio = 1, legend.position = "bottom") + 
  scale_color_botanical(palette = "fern")  
```

```{r fig.height= 8}
set.seed(123)
seed <- sample.int(10000, 20)
# kol_1d_2var_better_random_sim <- list()
# kol_1d_2var_better_random_sim_polish <- list()
# for (i in 1: length(seed)){
#   cat("i = ", i, "seed = ", seed[i], "\n")
#   set.seed(seed[i])
#   kol_1d_2var_better_random_sim[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_better, max.tries = 500, t0 = 1e-6, alpha = 0.7),
#                  rescale = FALSE, verbose = TRUE)
# 
# 
#   last_basis <- get_best(kol_1d_2var_better_random_sim[[i]])$basis %>% .[[1]]
# 
#   set.seed(seed[i])
#   kol_1d_2var_better_random_sim_polish[[i]] <-
#     animate_dist(boa6,
#                  tour_path = guided_tour(norm_kol(), d = 1,
#                                          search_f = search_polish, max.tries = 200),
#                  rescale = FALSE, verbose = TRUE,
#                  start = last_basis)
#   
# }

#save(kol_1d_2var_better_random_sim, file = here::here("data", "kol_1d_2var_better_random_sim.rda"))
#save(kol_1d_2var_better_random_sim_polish, file = here::here("data", "kol_1d_2var_better_random_sim_polish.rda"))

@@

all_summary <- best$basis %>% 
  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
  as_tibble() %>% 
  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
  mutate(IndexVal = best$index_val, seed = as.factor(best$seed)) %>% 
  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
  group_by(seed, IndexVal) %>% 
  filter(abs(value) == max(abs(value))) %>% 
  arrange(IndexVal) %>% ungroup()

@@

pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
  group_by(seed) %>% arrange(IndexVal) %>% 
  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
  geom_point(data = random, col = "grey", size = 0.3) +
  geom_point(aes(col = VarFound), alpha = 0.05) + 
  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
  geom_point(data = pca$aug %>% get_best(group = seed), size = 5) +
  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
  theme_void() + 
  theme(aspect.ratio = 1, legend.position = "bottom") + 
  scale_color_botanical(palette = "fern") 
```

@@

dt <- bind_rows(all_tidy_better %>% filter(seed == 1842) %>% mutate(algo = "better V2"),
          all_tidy_better %>% filter(seed == 9982) %>% mutate(algo = "better V7"),
          all_tidy %>% filter(seed == 4761) %>% mutate(algo = "better random"))

@@

p2 <- pca$aug %>% filter(method != "randomly_generated" & info != "theoretical") %>% 
  ggplot(aes(x = PC1, y = PC2, col = algo)) + 
  geom_point(data = random, col = "grey") +
  geom_point(aes(col = algo), alpha = 0.05) + 
  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
  geom_point(data = pca$aug %>% get_best(group = algo), size = 5) +
  geom_point(data = pca$aug %>% get_start() %>% filter(method %in% c("search_better_random", "search_better")),  size = 3) +
  facet_wrap(vars(algo)) +
  theme_void() + 
  theme(legend.position = "none", aspect.ratio = 1) + 
  scale_color_botanical(palette = "banksia")
